Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-12-02 analysis/paper/paper.qmd
Edited plots

library(cowplot)
plot_grid(plot_mass_cluster + guides(shape = "none"),
          plot_tcsa_mass_cluster,
          labels = "AUTO",
          rel_widths = c(0.5, 1))

@@

```{r}
#| label: fig-tcsa-size
#| fig-cap: "Artifact size and TSCA values by lithic raw material"
#| fig-width:   3.5
#| fig-height:  3

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point() +
    labs(x = "Weight (g)", 
       y = "TCSA",
       colour='Raw material') +
  geom_smooth(method=lm, se=FALSE, size = 0.4)+
  theme_bw(base_size = 6) +
  theme(
  legend.title = element_text(size = 5),
  legend.text = element_text(size = 4),
  legend.position = c(.8,.3)
  )
```

@@

```{r}
#| label: fig-tcsa-radio-carbon
#| fig-cap: "A: Distribution of TCSA values over time. The light gray shade indicates Marine isotope stages (MIS) 3, the gray shade indicates MIS 2 and the darkest shade indicates the duration of LGM. B: Distribution of TCSA values grouped by LGM event"
#| fig-width:   3.8
#| fig-height:  4

temporal_tcsa <- ggplot(tcsa_site_ass, 
       aes(y = min_tcsa,
           x = reorder(label_site, -age_ka))) +
   geom_boxplot() +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "") +
  theme(axis.text.x = element_text(size = 6)) +
  scale_x_discrete(guide = guide_axis(angle = 45))

@@

all_time <- temporal_tcsa +
 annotate("rect", 
           xmin = 0,
           xmax = 12.2,
           ymin = 0,
           ymax = 300,
           fill = "lightgray",
           alpha = 0.2) +
  annotate("text", 
               size = 3.5,
               label = "MIS 3",
               hjust = 0,
           x = 5,
           y = 260) +
   annotate("rect", 
           xmin = 12.2,
           xmax = 24.5,
           ymin = 0,
           ymax = 300,
           alpha = 0.2) +
  annotate("text", 
               size = 3.5,
               label = "MIS 2",
               hjust = 0,
           x = 17,
           y = 260) +
   annotate("rect", 
           xmin = 13.5,
           xmax = 21,
           ymin = 0,
           ymax = 230,
           alpha = 0.2) +
  annotate("text", 
               size = 2.3,
               label = "LGM",
               hjust = 0,
           x = 17.3,
           y = 210)

@@

# box plot for each MIS stage
three_time <-  ggplot(tcsa_site_ass_mis, 
       aes(y = min_tcsa,
           x = factor(mis_stage, level = c('After', 'LGM', 'Before')))) +
   geom_boxplot() +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "Time Period") +
  geom_quasirandom(alpha = 0.4,
                   size = 1.9) +
    scale_x_discrete(limits=rev) +
     annotate("rect", 
           xmin = 0.3,
           xmax = 1,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.4,
      #      y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 1,
           xmax = 1.7,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Un-poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.5,
      #      y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 1.7,
           xmax = 2.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Dart tips",
      #          hjust = 0,
      #      x = 3.6,
      #      y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 2.4,
           xmax = 3.1,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Javelin tips",
      #          hjust = 0,
      #      x = 3.7,
      #      y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 3.1,
           xmax = 3.8,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) 
# + annotate("text", 
#                label = "Stabbing\nspear tips",
#                size = point_function_group_label_size,
           #     hjust = 0,
           #     lineheight = 0.75,
           # x = 3.8,
           # y = (200-80) /2 + 80 )

plot_grid(all_time, 
          three_time, 
          labels = c('A', 'B'),
          ncol = 1,
          rel_heights = c(2.7, 1.5))

@@

```{r}
#| label: fig-tcsa-per-sites
#| fig-cap: "TCSA values by archaeological site."
#| fig-width:   4.5
#| fig-height:  4

 ggplot(tcsa_site,
              aes(y = min_tcsa, 
                  x = label_site)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 6)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 2.3)+
   annotate("rect", 
           xmin = 7.4,
           xmax = 7.5,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 7.4,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 7.5,
           xmax = 7.6,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 7.5,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 7.6,
           xmax = 7.7,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = 7.6,
           y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 7.7,
           xmax = 7.8,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = 7.7,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 7.8,
           xmax = 7.9,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing\nspear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = 7.8,
           y = (200-80) /2 + 80 )

@@

```{r}
#| label: fig-gis-tcsa-sites
#| fig-cap: "TCSA values by assemblages and ecoregions. A: Modern vegetation map defined by temporature and precipitation. The individual points indicate assemblages. B: TCSA distribution for vegitation zones. The individual points indicate artifacts. C: Tukey test result for vegitation zones. D: Ecoregion map based on geographical boundaries and ecological conditions (modified the map from @lee2008relationship). E: TCSA distribution for ecoregion zones, excluding zones with no artifacts. F: Tukey test result for ecoregion zones"
#| fig-width:   4.3
#| fig-height:  5.8

@@

eco_plot <- ggplot() +
  geom_sf(data = veg_all, aes(fill = "grey")) +
   geom_sf_text(data = veg_all, aes(label = ecoregion), size =1.5)+
  geom_sf(data = pts_all, aes(color = min_tcsa)) +
  theme_minimal() +
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, size = 6))+
  theme(axis.text.y = element_text(size = 6))+
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
  theme(legend.position = "none")

eco_box <- ggplot(points_in_veg_zones_all) +
  aes(ecoregion,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Vegetation") + 
  ylab("TCSA") +
  theme_minimal()

@@

for_plot <- ggplot() +
  geom_sf(data = veg_forest, aes(fill = "grey")) +
  geom_sf_text(data = veg_forest, aes(label = Zone), size =1.5)+
  geom_sf(data = pts_forest, aes(color = min_tcsa)) +
  theme_minimal()+
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, size = 6)) +
  theme(axis.text.y = element_text(size = 6)) +
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank())

@@

for_box <- ggplot(points_in_veg_zones_forest_box) +
  aes(Zone,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Ecoregion") + 
  ylab("TCSA") +
  theme_minimal()

@@

# plot Tukey HSD
eco_tur_box <- 
eco_tur$ecoregion %>% 
  as_tibble(rownames = "zones") %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones)) +
  geom_point(aes(diff,
                 zones),
             size = 4) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\nvegetation zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8)

@@

# plot Tukey HSD
for_tur_box <- 
for_tur$`as.factor(Zone)` %>% 
  as_tibble(rownames = "zones") %>% 
  mutate(zones = fct_reorder(zones, diff)) %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones)) +
  geom_point(aes(diff,
                 zones),
             size = 4) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\necological zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8)

plot_grid(eco_plot, 
          for_plot, 
          eco_box, 
          for_box, 
          eco_tur_box, 
          for_tur_box,
          labels = c('A', 'D', 'B', 'E', 'C', 'F'),
          ncol = 2)

>>>>

library(cowplot)
plot_grid(plot_mass_cluster + guides(shape = "none"),
          plot_tcsa_mass_cluster,
          labels = "AUTO",
          label_size = 9,
          rel_widths = c(0.5, 1))

@@

```{r}
#| label: fig-tcsa-size
#| fig-cap: "Artifact size and TSCA values by lithic raw material"
#| fig-height:  3.5

ggplot(tcsa_mass_no_na,
       aes(`Weight(g)`,
           min_tcsa,
           colour = raw_materials_top)) +
  geom_point(size = 1) +
    labs(x = "Weight (g)", 
       y = "TCSA",
       colour='Raw material') +
  geom_smooth(method=lm, se=FALSE, size = 0.3)+
  theme_bw(base_size = 6) +
  theme(
  legend.title = element_text(size = 4),
  legend.text = element_text(size = 3),
  legend.position = c(.8,.3),
  legend.key.height = unit(4.5, 'mm'),
  legend.key.width = unit(2, 'mm'))

@@

```{r}
#| label: fig-tcsa-radio-carbon
#| fig-cap: "A: Distribution of TCSA values over time. The light gray shade indicates Marine isotope stages (MIS) 3, the gray shade indicates MIS 2 and the darkest shade indicates the duration of LGM. B: Distribution of TCSA values grouped by LGM event"
#| fig-height:  5

temporal_tcsa <- ggplot(tcsa_site_ass, 
       aes(y = min_tcsa,
           x = reorder(label_site, -age_ka))) +
   geom_boxplot(size = 0.2) +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "") +
  theme(axis.text.x = element_text(size = 3.5),
        axis.text.y = element_text(size = 3.5),
        text = element_text(size=5)) +
  scale_x_discrete(guide = guide_axis(angle = 45))

@@

all_time <- temporal_tcsa +
 annotate("rect", 
           xmin = 0,
           xmax = 12.2,
           ymin = 0,
           ymax = 300,
           fill = "lightgray",
           alpha = 0.2) +
  annotate("text", 
               size = 3,
               label = "MIS 3",
               hjust = 0,
           x = 5,
           y = 260) +
   annotate("rect", 
           xmin = 12.2,
           xmax = 24.5,
           ymin = 0,
           ymax = 300,
           alpha = 0.2) +
  annotate("text", 
               size = 3,
               label = "MIS 2",
               hjust = 0,
           x = 17,
           y = 260) +
   annotate("rect", 
           xmin = 13.5,
           xmax = 21,
           ymin = 0,
           ymax = 230,
           alpha = 0.2) +
  annotate("text", 
               size = 2.1,
               label = "LGM",
               hjust = 0,
           x = 17.3,
           y = 210)

@@

# box plot for each MIS stage
three_time <-  ggplot(tcsa_site_ass_mis, 
       aes(y = min_tcsa,
           x = factor(mis_stage, level = c('After', 'LGM', 'Before')))) +
   geom_boxplot(size = 0.4) +
   theme_minimal() +
   theme(legend.position="none", 
         axis.text.x = element_text(size = 3.5),
        axis.text.y = element_text(size = 3.5),
        text = element_text(size=5)) +
    labs(y = "TCSA", 
        x = "Time Period") +
  geom_quasirandom(alpha = 0.4,
                   size = 1.1) +
    scale_x_discrete(limits=rev) +
     annotate("rect", 
           xmin = 0.3,
           xmax = 1,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.4,
      #      y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 1,
           xmax = 1.7,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Un-poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.5,
      #      y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 1.7,
           xmax = 2.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Dart tips",
      #          hjust = 0,
      #      x = 3.6,
      #      y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 2.4,
           xmax = 3.1,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Javelin tips",
      #          hjust = 0,
      #      x = 3.7,
      #      y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 3.1,
           xmax = 3.8,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) 
# + annotate("text", 
#                label = "Stabbing\nspear tips",
#                size = point_function_group_label_size,
           #     hjust = 0,
           #     lineheight = 0.75,
           # x = 3.8,
           # y = (200-80) /2 + 80 )

plot_grid(all_time, 
          three_time, 
          labels = "AUTO",
          label_size = 9,
          ncol = 1,
          rel_heights = c(2.7, 2))

@@

```{r}
#| label: fig-tcsa-per-sites
#| fig-cap: "TCSA values by archaeological site."
#| fig-height:  3

 ggplot(tcsa_site,
              aes(y = min_tcsa, 
                  x = label_site)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 5)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 1.6)+
   annotate("rect", 
           xmin = 0.4,
           xmax = 1.9,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 1.9,
           xmax = 3.4,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 3.4,
           xmax = 4.9,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 4.9,
           xmax = 6.4,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
     annotate("rect", 
           xmin = 6.4,
           xmax = 7.9,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2)

@@

```{r}
#| label: fig-gis-tcsa-sites
#| fig-cap: "TCSA values by assemblages and ecoregions. A: Modern vegetation map defined by temporature and precipitation. The individual points indicate assemblages. B: TCSA distribution for vegitation zones. The individual points indicate artifacts. C: Tukey test result for vegitation zones. D: Ecoregion map based on geographical boundaries and ecological conditions (modified the map from @lee2008relationship). E: TCSA distribution for ecoregion zones, excluding zones with no artifacts. F: Tukey test result for ecoregion zones"
#| fig-height:  5

@@

eco_plot <- ggplot() +
  geom_sf(data = veg_all, aes(fill = "grey")) +
   geom_sf_text(data = veg_all, aes(label = ecoregion), size =1.5)+
  geom_sf(data = pts_all, aes(color = min_tcsa), size = 1.1) +
  theme_minimal() +
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, size = 4.5))+
  theme(axis.text.y = element_text(size = 4.5))+
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
  theme(legend.position = "none")

eco_box <- ggplot(points_in_veg_zones_all) +
  aes(ecoregion,
      min_tcsa) +
  geom_boxplot(size = 0.2) +
  geom_quasirandom(alpha = 0.3,
                   size = 1.3) +
  xlab("Vegetation") + 
  ylab("TCSA") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

@@

for_plot <- ggplot() +
  geom_sf(data = veg_forest, aes(fill = "grey")) +
  geom_sf_text(data = veg_forest, aes(label = Zone), size =1.5)+
  geom_sf(data = pts_forest, aes(color = min_tcsa), size = 1.1) +
  theme_minimal()+
  scale_fill_brewer(palette="Greys") +
  labs(color="TCSA") +
  guides(fill = FALSE) +
  theme(axis.text.x = element_text(angle = 45, size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
  theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
  theme(legend.title = element_text(size= 6),
        legend.key.height = unit(4.5, 'mm'),
        legend.key.width = unit(2, 'mm'),
        legend.text = element_text(size = 4))

@@

for_box <- ggplot(points_in_veg_zones_forest_box) +
  aes(Zone,
      min_tcsa) +
  geom_boxplot(size = 0.2) +
  geom_quasirandom(alpha = 0.3,
                   size = 1.3) +
  xlab("Ecoregion") + 
  ylab("TCSA") +
  theme_minimal() +
      theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

@@

# plot Tukey HSD
eco_tur_box <- 
eco_tur$ecoregion %>% 
  as_tibble(rownames = "zones") %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones),
                 size = 0.4) +
  geom_point(aes(diff,
                 zones),
             size = 1) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\nvegetation zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

@@

# plot Tukey HSD
for_tur_box <- 
for_tur$`as.factor(Zone)` %>% 
  as_tibble(rownames = "zones") %>% 
  mutate(zones = fct_reorder(zones, diff)) %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones), 
                 size = 0.4) +
  geom_point(aes(diff,
                 zones),
             size = 1) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\necological zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(size = 4.5)) +
  theme(axis.text.y = element_text(size = 4.5)) +
   theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6))

plot_grid(eco_plot, 
          for_plot, 
          eco_box, 
          for_box, 
          eco_tur_box, 
          for_tur_box,
          labels = c('A', 'D', 'B', 'E', 'C', 'F'),
          label_size = 9,
          ncol = 2)
